---
title: "Project In Statistics"
author: "Anders Gantzhorn Kristensen"
date: "2023-04-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(momentuHMM)
theme_set(theme_bw())
```

# Test

Check if everything works

```{r, include=FALSE}
eagleDataOriginal <- readr::read_csv("GoldenEagle_FinalData_Dryad_FE.csv")
eagleData <- eagleDataOriginal
eagleData %>% mutate(Segment_ID = factor(Segment_ID))
eagleData <- eagleData %>% group_by(Segment_ID) %>%
    mutate(vertical_steps = Altitude - lag(Altitude, default = first(Altitude))) %>% 
    rename(horizontal_steps = Step_length) %>% 
    slice(2:n())
```


# Estimation using Weibull- and Gamma distributions
First assume that Step_length is Weibull distributed. The MLE of $k$ is the solution to
$$
0 = \frac{\sum_{i = 1}^N X_i^k\log(X_i)}{\sum_{i = 1}^N X_i^k} - \frac{1}{k} - \frac{1}{N}\sum_{i = 1}^N \log(X_i)
$$

whereas the MLE of $\lambda$ given $k$ is:
$$
\hat{\lambda} = \left(\frac{1}{N} \sum_{i = 1}^N X_i^k\right)^{\frac{1}{k}}
$$
Thus we can calculate the MLE assuming step-length has a Weibull distribtion
```{r}
k_hat <- uniroot(function(k) {sum(eagleData$horizontal_steps^k * log(eagleData$horizontal_steps)) / sum(eagleData$horizontal_steps^k) - 1/k - mean(log(eagleData$horizontal_steps))},
                    c(0.1, 10))$root
lambda_hat <- (mean(eagleData$horizontal_steps^k_hat))^(1/k_hat)
```

We can simulate from this distribution and compare with the data

```{r}
simWei <- rweibull(50000, shape = k_hat, scale = lambda_hat)

tibble(x = c(simWei, eagleData$horizontal_steps), type = rep(c("sim", "obs"), times = c(length(simWei), length(eagleData$horizontal_steps)))) %>% ggplot(aes(x = x, color = type)) + geom_density()
```
Might not be the best fit to the Data, but it is a start.\
<!-- Try to fit Gamma distribution:\ -->
<!-- MLE of shape parameter in gamma distribution is the solution to the equation: -->
<!-- $$ -->
<!-- \log(\alpha)-\psi(\alpha) = \log\left(\frac{1}{N}\sum_{i=1}^N\log(X_i)\right)-\frac{1}{N}\sum_{i = 1}^N\log(X_i) -->
<!-- $$ -->
<!-- Whereas the rate parameter is given by -->
<!-- $$\hat{\beta} = \frac{\hat{\alpha}}{\sum_{i = 1}^N X_i}$$ -->

<!-- ```{r} -->
<!-- alpha_hat <- uniroot(function(alpha) {log(alpha)-digamma(alpha)-log(mean(eagleData$horizontal_steps))+mean(log(eagleData$horizontal_steps))},c(0.1, 10))$root -->
<!-- beta_hat <- alpha_hat/mean(eagleData$horizontal_steps) -->

<!-- simGamma <- rgamma(50000, shape = alpha_hat, rate = beta_hat) -->

<!-- tibble(x = c(simGamma, eagleData$horizontal_steps), type = rep(c("sim", "obs"), times = c(length(simGamma), length(eagleData$horizontal_steps)))) %>% ggplot(aes(x = x, color = type)) + geom_density() -->
<!-- ``` -->
<!-- Indicating a bit better of a fit maybe. -->

# Fitting the vertical steps to a normal distribution
The MLE of the mean in the normal distribution is:
$$
\hat{\mu} = \frac{1}{N}\sum_{i = 1}^N X_i
$$
Whereas the variance is:
$$
\hat{\mu} = \frac{1}{N}\sum_{i = 1}^N \left(X_i-\hat{\mu}\right)^2
$$
Which we implement.
```{r}
mu_hat <- mean(eagleData$vertical_steps)
sigmasq_hat <- var(eagleData$vertical_steps)

```


```{r}
prepEagleData <- momentuHMM::prepData(data = data.frame(ID = eagleData$Segment_ID,
                                       vertical_steps = eagleData$vertical_steps,
                                       horizontal_steps = eagleData$horizontal_steps),
                                       coordNames = NULL)
```
# We fit the model using the parameters from above and decode the states
```{r}
N <- 3 
# Initial values for weibull distributions of horizontal distances
k0 <- rep(k_hat, times = N)
lambda0 <- rep(lambda_hat, times = N)

# Initial values for normal distributions for vertical distances
mu0 <- rep(mu_hat, times = N)
sigma0 <- rep(sigmasq_hat, times = N)
eagleFit1 <- momentuHMM::fitHMM(prepEagleData,
                   nbStates = N,
                   dist = list(horizontal_steps = "weibull",
                          vertical_steps = "norm"),
                   Par0 = list(horizontal_steps = c(k0, lambda0),
                          vertical_steps = c(mu0, sigma0))
)

decodedStatesFit1 <- viterbi(m = eagleFit1)

```

```{r}
eagleData <- tibble(eagleData, decodedstates = decodedStatesFit1)
```

```{r}
eagleData <- eagleData %>% mutate(decodedstates = factor(decodedstates))

ggplot(filter(eagleData), aes(x = vertical_steps, y = log(horizontal_steps), col = decodedstates)) + geom_point()
```

