---
title: "Project In Statistics"
author: "Anders Gantzhorn Kristensen"
date: "2023-04-29"
output: html_document
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(momentuHMM)
library(maps)
library(mapdata)
library(ggmap)
library(mapview)
library(ggthemes)
theme_set(theme_bw())
proj_palette <- c("#E69F00", "#56B4E9", "#009E73",
                         "#F0E442", "#0072B2", "#D55E00",
                         "#CC79A7")
```

# Test

Check if everything works

```{r, include=FALSE, message = FALSE}
eagleDataOriginal <- readr::read_csv("Eagle_data_with_GPS_positions.csv")
eagleData <- eagleDataOriginal
eagleData %>% mutate(Segment_ID = factor(Segment_ID))
eagleData <- eagleData %>% group_by(Segment_ID) %>%
    mutate(vertical_steps = Altitude - lag(Altitude, default = first(Altitude))) %>% 
    rename(horizontal_steps = Step_length) %>% 
    slice(2:n())
```

# We need to try to identify three distinct states.

# Estimation using Weibull- and Gamma distributions
First assume that Step_length is Weibull distributed. The MLE of $k$ is the solution to
$$
0 = \frac{\sum_{i = 1}^N X_i^k\log(X_i)}{\sum_{i = 1}^N X_i^k} - \frac{1}{k} - \frac{1}{N}\sum_{i = 1}^N \log(X_i)
$$

whereas the MLE of $\lambda$ given $k$ is:
$$
\hat{\lambda} = \left(\frac{1}{N} \sum_{i = 1}^N X_i^k\right)^{\frac{1}{k}}
$$
Thus we can calculate the MLE assuming step-length has a Weibull distribtion
```{r message = FALSE}
k_hat <- uniroot(function(k) {sum(eagleData$horizontal_steps^k * log(eagleData$horizontal_steps)) / sum(eagleData$horizontal_steps^k) - 1/k - mean(log(eagleData$horizontal_steps))},
                    c(0.1, 10))$root
lambda_hat <- (mean(eagleData$horizontal_steps^k_hat))^(1/k_hat)
```

We can simulate from this distribution and compare with the data

```{r message = FALSE}
simWei <- rweibull(50000, shape = k_hat, scale = lambda_hat)

tibble(x = c(simWei, eagleData$horizontal_steps), type = rep(c("sim", "obs"), times = c(length(simWei), length(eagleData$horizontal_steps)))) %>% ggplot(aes(x = x, color = type)) + geom_density()
```
Might not be the best fit to the Data, but it is a start.\
<!-- Try to fit Gamma distribution:\ -->
<!-- MLE of shape parameter in gamma distribution is the solution to the equation: -->
<!-- $$ -->
<!-- \log(\alpha)-\psi(\alpha) = \log\left(\frac{1}{N}\sum_{i=1}^N\log(X_i)\right)-\frac{1}{N}\sum_{i = 1}^N\log(X_i) -->
<!-- $$ -->
<!-- Whereas the rate parameter is given by -->
<!-- $$\hat{\beta} = \frac{\hat{\alpha}}{\sum_{i = 1}^N X_i}$$ -->

```{r}
alpha_hat <- uniroot(function(alpha) {log(alpha)-digamma(alpha)-log(mean(eagleData$horizontal_steps))+mean(log(eagleData$horizontal_steps))},c(0.1, 10))$root
beta_hat <- alpha_hat/mean(eagleData$horizontal_steps)

simGamma <- rgamma(50000, shape = alpha_hat, rate = beta_hat)

tibble(x = c(simGamma, eagleData$horizontal_steps), type = rep(c("sim", "obs"), times = c(length(simGamma), length(eagleData$horizontal_steps)))) %>% ggplot(aes(x = x, color = type)) + geom_density()
```
Indicating a bit better of a fit maybe.

# Fitting the vertical steps to a normal distribution
The MLE of the mean in the normal distribution is:
$$
\hat{\mu} = \frac{1}{N}\sum_{i = 1}^N X_i
$$
Whereas the variance is:
$$
\hat{\mu} = \frac{1}{N}\sum_{i = 1}^N \left(X_i-\hat{\mu}\right)^2
$$
Which we implement.
```{r message = FALSE}
mu_hat <- mean(eagleData$vertical_steps)
sigmasq_hat <- var(eagleData$vertical_steps)

```


```{r message = FALSE}
prepEagleData <- momentuHMM::prepData(data = data.frame(ID = eagleData$Segment_ID,
                                       vertical_steps = eagleData$vertical_steps,
                                       horizontal_steps = eagleData$horizontal_steps),
                                       coordNames = NULL)
```
We also need the number of observations per animal for the validation process later.
```{r message = FALSE}
obsPerAnimal <- as.list(as.numeric(table(prepEagleData$ID))) 
```

# Model fitting

We fit the model using the parameters from above and decode the states
```{r message = FALSE}
N <- 3 
# Initial values for weibull distributions of horizontal distances
k0 <- rep(k_hat, times = N)
lambda0 <- rep(lambda_hat, times = N)

# Initial values for normal distributions for vertical distances
mu0 <- rep(mu_hat, times = N)
sigma0 <- rep(sigmasq_hat, times = N)
eagleFit1 <- momentuHMM::fitHMM(prepEagleData,
                   nbStates = N,
                   dist = list(horizontal_steps = "weibull",
                          vertical_steps = "norm"),
                   Par0 = list(horizontal_steps = c(k0, lambda0),
                          vertical_steps = c(mu0, sigma0))
)

decodedStatesFit1 <- viterbi(m = eagleFit1)

```
Use the generic plot method on the model
```{r message = FALSE}
#plot(eagleFit1)
```

```{r message = FALSE}
numberAfterState <- decodedStatesFit1 %>% tibble(stateNum=.) %>% group_by(stateNum) %>% count() %>% ungroup() %>% mutate(prop = n/sum(n))
```

```{r message = FALSE}
# Set the parameters for the Weibull distributions
shape_params <- eagleFit1$mle$horizontal_steps[1,]
scale_params <- eagleFit1$mle$horizontal_steps[2,]

# Set the number of observations for each distribution
num_observations <- c(numberAfterState$n[numberAfterState$stateNum == 1], 
                      numberAfterState$n[numberAfterState$stateNum == 2], 
                      numberAfterState$n[numberAfterState$stateNum == 3])

# Generate the x values
x <- density(eagleData$horizontal_steps)$x

# Calculate the density values for each distribution
pdf_values <- sapply(1:N, function(i) {
  density <- dweibull(x, shape = shape_params[i], scale = scale_params[i])
  density <- density * numberAfterState$prop[numberAfterState$stateNum == i]
  density
})



# Create a data frame for the density values
df <- data.frame(x = rep(x, (N+1)),
                 density = as.vector(cbind(pdf_values, rowSums(pdf_values))),
                 state = factor(rep(c(1:N, "Total"), each = length(x))))

# Plot the density functions using ggplot2
ggplot() + 
  geom_histogram(data = eagleData, aes(x = horizontal_steps, after_stat(density)), fill = "gray", col = "black") +
  geom_line(data = filter(df, state != "Total"), aes(x = x, y = density, color = state),
                lwd = 1, alpha = 1.5) +
  geom_line(data = filter(df, state == "Total"), aes(x = x, y = density),
                lwd = 1, alpha = 1.5, linetype = "dashed") +
  xlab("x") +
  ylab("Density") +
  scale_color_manual(values = proj_palette)
```

Add decoded states to the dataset
```{r message = FALSE}
eagleData <- tibble(eagleData, decodedstates = decodedStatesFit1)
```
And plot the results
```{r message = FALSE}
eagleData <- eagleData %>% mutate(decodedstates = factor(decodedstates))

ggplot(eagleData, aes(x = log(horizontal_steps), y = vertical_steps)) + geom_point(alpha = .6) + facet_wrap(~decodedstates,ncol = 1)
```

# Assess the fitted model
Look at simulations from the model and whether they show similar patterns as the observed data.

```{r message = FALSE, include = FALSE}
eagleSim <- momentuHMM::simData(nbAnimals = max(as.numeric(levels(prepEagleData$ID))),
                                nbStates = N,
                                model = eagleFit1,
                                obsPerAnimal = obsPerAnimal)
eagleSim
```

```{r message = FALSE}
bind_rows(eagleSim, prepEagleData) %>%
    mutate(Data = rep(c("Sim", "Obs"), each = nrow(eagleSim))) %>%
  pivot_longer(cols = -c(ID, Data), names_to = "Movement_type", values_to = "Move") %>% 
  ggplot(aes(x = Move, fill = Data)) + geom_density() +
  scale_fill_manual(values = proj_palette) + facet_wrap(~Movement_type, ncol = 1, scale = "free")
```
# Assess fit

```{r message = FALSE}
pseudoReseagleFit1 <- momentuHMM::pseudoRes(eagleFit1)
pseudoResTibble <- tibble(horizontal_stepsRes = pseudoReseagleFit1[[1]],
                          vertical_stepsRes = pseudoReseagleFit1[[2]]
                          ) %>% mutate(time = row_number())
```

```{r message = FALSE}
# Chain independence:
pseudoResTibble %>% ggplot(aes(x = time, y = horizontal_stepsRes)) +
  geom_line(color = proj_palette[5])
```

```{r message = FALSE}
# Histogram of residuals with standard normal distribution on top
pseudoResTibble %>% ggplot(aes(x = vertical_stepsRes)) +
  geom_histogram(aes(y = after_stat(density)),col = "black", fill = proj_palette[6], alpha = 0.8) + 
  stat_function(fun = dnorm, args = list(mean = 0, sd = 1), color = proj_palette[5], linewidth = 1)
```

```{r message = FALSE}
# QQ-plot of residuals
pseudoResTibble %>% ggplot(aes(sample = horizontal_stepsRes)) + 
  geom_qq() +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  labs(title = "QQ Plot of Horizontal Steps Pseudo Residuals")
```

```{r message = FALSE}
# Make autocorrelation plots
acf_vals <- acf(pseudoResTibble$vertical_stepsRes[is.finite(pseudoResTibble$vertical_stepsRes)], plot = FALSE)
acftib <- tibble(ACF = acf_vals$acf, lag = acf_vals$lag)
ggplot(acftib, aes(x = lag, y = ACF)) +
  geom_hline(yintercept = 0, color = "black") +
  geom_hline(yintercept = c(0.025, -0.025), linetype = "dashed", color = proj_palette[3]) +
  geom_segment(aes(xend = lag, yend = 0), color = "black") +
  xlab("Lag") +
  ylab("Autocorrelation")
```

```{r message = FALSE}
 # Find segment with the most observations:
favourite_bird <- eagleData %>% group_by(Segment_ID) %>% summarise(n = n()) %>% ungroup() %>% 
  slice_max(n) %>% .$Segment_ID

# Find interesting birds:
eagleData %>% group_by(Segment_ID, decodedstates) %>% summarise(n = n()) %>% ungroup() %>% group_by(Segment_ID) %>% mutate(prop = n/sum(n)) %>% select(Segment_ID, decodedstates, prop) %>%  pivot_wider(names_from = "decodedstates", values_from = "prop") %>%
  mutate(across(where(is.double), ~ replace_na(., 0))) %>% filter(`1`>0.25 & `2`>0.25 & `3`>0.25)


favourite_birdTibble <- eagleData %>% filter(Segment_ID == 48)
mapview::mapview(favourite_birdTibble, xcol = "Longitude", ycol = "Latitude", crs = 4326,
                 grid = FALSE, zcol = "decodedstates")
```
Try plotting with a different package
```{r}
favoriteBirdMap <- ggmap::get_stamenmap(bbox = c(left = min(favourite_birdTibble$Longitude),
                              right = max(favourite_birdTibble$Longitude),
                              bottom = min(favourite_birdTibble$Latitude),
                              top = max(favourite_birdTibble$Latitude)),
                              maptype = "terrain",
                              crop = FALSE,
                              zoom = 11)
allBirdsMap <- ggmap::get_stamenmap(bbox = c(left = min(eagleData$Longitude),
                              right = -73,
                              bottom = min(eagleData$Latitude),
                              top = 43),
                              maptype = "terrain-background",
                              crop = FALSE,
                              zoom = 8)

ggmap(favoriteBirdMap)+
  geom_line(data = favourite_birdTibble, aes(x = Longitude, y = Latitude),
            lwd = 0.1, col = "black") + 
  geom_point(data = favourite_birdTibble, aes(x = Longitude, y = Latitude, col = decodedstates)) +
  scale_color_manual(values = proj_palette) +
  labs(title = "Segment 48 path",
       x = "Longitude",
       y = "Latitude") +
  theme(plot.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        legend.title = element_text(size = 12, face = "bold"),
        legend.text = element_text(size = 10),
        legend.position = "bottom")
```

```{r}
ggmap(allBirdsMap) +
  geom_point(data = eagleData, aes(x = Longitude, y = Latitude, group = Segment_ID),
            size = 0.0001, col = "black") + 
  labs(title = "Most birds",
       x = "Longitude",
       y = "Latitude") +
  theme(plot.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        legend.title = element_text(size = 12, face = "bold"),
        legend.text = element_text(size = 10),
        legend.position = "bottom")
```

